### 2019-04-29 report/overview.tex
translation at the kilometer

\chapter{Overview of Optimal Transport}

\myminitoc

Optimal transport aims to define a geometry and a distance on the set of
mesures of probability on a space.

\sect{Introduction}

\subs{\textsc{Monge}'s Problem}

\begin{center}
	\begin{tikzpicture}[>=stealth']
		\draw[fill=brown!20] (-1, -1.8) rectangle (9, 0);
		\draw[fill=red, domain=0:2, smooth, variable=\x, fill opacity=0.9]
			plot({1.5*\x}, {\x * (27.21 + \x * (-126.6 + \x * (229.9 + \x * (-200.7
				+ \x * (84.24 - 13.61 * \x)))))}) -- (3.05, 0);
		\node[red] at (-0.2, 0.7) {$\mu$};
		\draw[fill=blue, fill opacity=0.9]
			(5, 0) -- (5.1, -1.2) -- (5.5, -1.2) -- (5.6, -1) -- (5.9, -1) --
			(6, -1.2) -- (6.4, -1.2) -- (6.5, -1) -- (6.8, -1) -- (6.9, -1.2) --
			(7.3, -1.2) -- (7.4, -1) -- (7.7, -1) -- (7.8, 0);
		\node[blue] at (8, -1.2) {$\nu$};
		\draw[ultra thick, purple] (0.4, 0) -- (0.4, 1.68);
		\draw[ultra thick, purple] (6.1, 0) -- (6.1, -1.2);
		\node[red] (x) at (0.4, -0.15) {$\bullet$};
		\node[blue] (y) at (6.1, 0.15) {$\bullet$};
		\node[below] at (x) {$x$};
		\node[above] at (y) {$y = T(x)$};
		\coordinate (y) at (6, 0.1);
		\draw[->, ultra thick, greenTikz, bend right=30]
			(x) edge node[below, black] {${\color{greenTikz}c}(x, T(x))$} (y);
	\end{tikzpicture}
\end{center}

\paragraph{Monge}

We work on a space of probability $\Omega$. We consider two mesures of
probability $\mu$ and $\nu$ on $\mathcal{P}(\Omega)$. We also consider a cost function
$c : \Omega \times \Omega \rightarrow \R$. This function can for instance be the Euclidean distance. Monge's problem is therefore the following

$$ \int_{\Omega} c \left( x, T(x) \right) \mu(x) $$
\vspace{-2mm}
$$ \text{s.t.} \quad \forall A \in \mathcal{P}(\Omega), \, \nu(A) = \mu \left( T^{-1}(A) \right) $$

The function $T$ is called a transport map. It carries the information
regarding the displacement of each elementary unit of probability: the mass
located in $x$ is moved toward $T(x)$. The function $c$ thus allows to
quantify the cost of the displacement of one elementary unit of mass. We
want to minimize the average distance travelled by mass units during the
optimal transport.

\paragraph{Limitations}

The issue with this formulation of an optimal transport problem is that it
doesn't handle discrete probability mesures. Indeed, if $\mu$ has a Dirac
impulse bigger than every Dirac impulse of $\nu$, then the transport map
cannot be build.

\subs{Kantorovich's problem}

\begin{center}
	\begin{tikzpicture}[>={latex}, scale=1.2]
		\draw[->, greenTikz, ultra thick] (5.20, 3.65) -- node[black, above] {60} (3.96, 3.47);
		\fill[red] (5.20, 3.65) circle (0.136) node[white] {\tiny 60};
		\fill[blue] (3.65, 3.42) circle (0.313) node[white] {120};
		\draw[->, greenTikz, ultra thick] (6.69, 0.87) -- node[black, above right] {90} (5.21, 1.91);
		\fill[red] (6.69, 0.87) circle (0.221) node[white] {90};
		\fill[blue] (0.33, 0.18) circle (0.221) node[white] {90};
		\draw[->, greenTikz, ultra thick] (3.73, 0.65) -- node[black, left] {60} (3.66, 3.11);
		\draw[->, greenTikz, ultra thick] (3.73, 0.65) -- node[black, above] {90} (0.55, 0.21);
		\fill[red] (3.73, 0.65) circle (0.409) node[white] {150};
		\fill[blue] (5.03, 2.04) circle (0.221) node[white] {90};

		\node[red, below] at (5.2, 3.55) {\textbf{1}};
		\node[red, right=6] at (6.69, 0.87) {\textbf{2}};
		\node[red, below right=9] at (3.73, 0.65) {\textbf{3}};
		\node[blue, left=8] at (3.65, 3.42) {\textbf{A}};
		\node[blue, left=6] at (0.33, 0.18) {\textbf{B}};
		\node[blue, above=6] at (5.03, 2.04) {\textbf{C}};
	\end{tikzpicture}
\end{center}

\paragraph{Changement de plan de transport}
\paragraph{Change of transport plan}
The idea of Kantorovich is to change the function $T : \Omega \rightarrow
\Omega$ into a probability $P$ on the product space $\mathcal{P}(\Omega \times
\Omega)$. On the upper example, Monge's formulation forbids to separate the
masse of size 150 into two separate masses. With Kantorovich's formulation,
this operation is accessible: the mass of $150$ is splitted into a mass of
$90$ and a mass of $60$, both sent to different places. In the discrete
case, we can represent what happens with a matrix.

\begin{center}
	\begin{tikzpicture}
		\draw[step=1, thick] (0, 0) grid (3, 3);
		\node at (0.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}A}}$};
		\node at (1.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}B}}$};
		\node at (2.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}C}}$};
		\node at (0.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}A}}$};
		\node at (1.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}B}}$};
		\node at (2.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}C}}$};
		\node at (0.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}A}}$};
		\node at (1.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}B}}$};
		\node at (2.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}C}}$};

		\node at (1.5, 3.5) {\large Transport Map};
		\node[blue] at (0.5, -0.5) {\Large 120};
		\node[blue] at (1.5, -0.5) {\Large 90};
		\node[blue] at (2.5, -0.5) {\Large 90};
		\node[red] at (-0.5, 0.5) {\Large 60};
		\node[red] at (-0.5, 1.5) {\Large 90};
		\node[red] at (-0.5, 2.5) {\Large 150};

		\node at (7.5, 3.5) {\large Distance matrix};
		\draw[step=1, thick] (6, 0) grid (9, 3);
		\node at (6.5, 0.5) {\Large $d_{{\color{red}3}{\color{blue}A}}$};
		\node at (7.5, 0.5) {\Large $d_{{\color{red}3}{\color{blue}B}}$};
		\node at (8.5, 0.5) {\Large $d_{{\color{red}3}{\color{blue}C}}$};
		\node at (6.5, 1.5) {\Large $d_{{\color{red}2}{\color{blue}A}}$};
		\node at (7.5, 1.5) {\Large $d_{{\color{red}2}{\color{blue}B}}$};
		\node at (8.5, 1.5) {\Large $d_{{\color{red}2}{\color{blue}C}}$};
		\node at (6.5, 2.5) {\Large $d_{{\color{red}1}{\color{blue}A}}$};
		\node at (7.5, 2.5) {\Large $d_{{\color{red}1}{\color{blue}B}}$};
		\node at (8.5, 2.5) {\Large $d_{{\color{red}1}{\color{blue}C}}$};
	\end{tikzpicture}
\end{center}

\paragraph{Discrete Case}

If we note $D$ our disance matrix, we obtain the following formulation in the discrete case:

$$ \min_P \langle P, D \rangle = \sum_i \sum_j p_{ij} d_{ij} $$
\vspace{-4mm}
$$ \text{s.t.} \quad \left\{ \begin{array}{ll}
\forall i & \sum_j p_{ij} = \mu_i \\
\forall j & \sum_i p_{ij} = \nu_j
\end{array} \right. $$

\paragraph{Continuous Case}

In the continuous setting, we introduce the following set of joint probabilities:

$$ \Pi(\mu, \nu) = \left\{ P \in \mathcal{P}(\Omega \times \Omega)~|~\forall A, B \in \Omega, \, P(A \times \Omega) = \mu(A), \, P(\Omega \times B) = \nu(B) \right\} $$

This set encodes the conditions of the discrete case but in the very
continuous setting. We consider once again our cost function $c$ defined
in the Monge's problem. Kantorovich's formulation now becomes the following
optimization problem:

$$ \inf_{P \in \Pi(\mu, \nu)} \E_P \left[ c(X, Y) \right] = \int \int c(x, y) P(dx, dy) $$
\subs{Wasserstein distance}

This optimization problem allows us to define a distance between probability mesures:

\DEF{
    Given a probability space $\Omega$, a cost function $c : \Omega \times
    \Omega \rightarrow \R$ and two probabilitiy mesures $\mu$ and $\nu$ in
    $\mathcal{P}(\Omega)$, the \textbf{p-Wasserstein-distance} between $\mu$
    and $\nu$ is defined by :
	$$ W_p(\mu, \nu) = \left( \inf_{P \in \Pi(\mu, \nu)} \int \int c(x,
	y)^p P(dx, dy) \right)^{1 / p} $$
	\vspace{-3mm}
}

This distance allows to compute barycenters between several probability
distributions, and obtain results that are way more natural than barycenters
obtained with other norms, like $l_2$ for instance. The following
figure compares an interpolation (progressive weighted barycenter of two
distributions) obtained via Wasserstein-distance and a linear interpolation.

\begin{center}
	\begin{tikzpicture}[scale=0.85]
		\node at (5.5, 4.5) {\large Interpolation linéaire};
		\node[red] at (0.5, 2.3) {$\mu$};
		\node[blue] at (9.5, 2.7) {$\nu$};
		\fill[red] (0.000, 0.000) -- (0.050, 0.403) -- (0.100, 0.752) -- (0.150, 1.051) -- (0.200, 1.302) -- (0.250, 1.509) -- (0.300, 1.676) -- (0.350, 1.806) -- (0.400, 1.902) -- (0.450, 1.967) -- (0.500, 2.004) -- (0.550, 2.016) -- (0.600, 2.006) -- (0.650, 1.976) -- (0.700, 1.929) -- (0.750, 1.868) -- (0.800, 1.794) -- (0.850, 1.710) -- (0.900, 1.618) -- (0.950, 1.521) -- (1.000, 1.419) -- (1.050, 1.315) -- (1.100, 1.211) -- (1.150, 1.108) -- (1.200, 1.008) -- (1.250, 0.912) -- (1.300, 0.822) -- (1.350, 0.738) -- (1.400, 0.662) -- (1.450, 0.595) -- (1.500, 0.537) -- (1.550, 0.490) -- (1.600, 0.455) -- (1.650, 0.431) -- (1.700, 0.419) -- (1.750, 0.420) -- (1.800, 0.435) -- (1.850, 0.462) -- (1.900, 0.503) -- (1.950, 0.556) -- (2.000, 0.623) -- (2.050, 0.703) -- (2.100, 0.796) -- (2.150, 0.900) -- (2.200, 1.017) -- (2.250, 1.144) -- (2.300, 1.281) -- (2.350, 1.428) -- (2.400, 1.584) -- (2.450, 1.746) -- (2.500, 1.915) -- (2.550, 2.089) -- (2.600, 2.267) -- (2.650, 2.447) -- (2.700, 2.627) -- (2.750, 2.807) -- (2.800, 2.984) -- (2.850, 3.156) -- (2.900, 3.321) -- (2.950, 3.478) -- (3.000, 3.624) -- (3.050, 3.757) -- (3.100, 3.874) -- (3.150, 3.973) -- (3.200, 4.052) -- (3.250, 4.107) -- (3.300, 4.136) -- (3.350, 4.137) -- (3.400, 4.105) -- (3.450, 4.038) -- (3.500, 3.933) -- (3.550, 3.787) -- (3.600, 3.595) -- (3.650, 3.355) -- (3.700, 3.063) -- (3.750, 2.715) -- (3.800, 2.307) -- (3.850, 1.836) -- (3.900, 1.297) -- (3.950, 0.686) -- (4.000, 0.000);
		\fill[blue] (7.000, 0.000) -- (7.050, 0.215) -- (7.100, 0.418) -- (7.150, 0.611) -- (7.200, 0.792) -- (7.250, 0.964) -- (7.300, 1.126) -- (7.350, 1.278) -- (7.400, 1.421) -- (7.450, 1.555) -- (7.500, 1.681) -- (7.550, 1.798) -- (7.600, 1.907) -- (7.650, 2.008) -- (7.700, 2.102) -- (7.750, 2.188) -- (7.800, 2.268) -- (7.850, 2.341) -- (7.900, 2.408) -- (7.950, 2.468) -- (8.000, 2.522) -- (8.050, 2.571) -- (8.100, 2.614) -- (8.150, 2.652) -- (8.200, 2.685) -- (8.250, 2.713) -- (8.300, 2.737) -- (8.350, 2.755) -- (8.400, 2.770) -- (8.450, 2.781) -- (8.500, 2.787) -- (8.550, 2.790) -- (8.600, 2.790) -- (8.650, 2.785) -- (8.700, 2.778) -- (8.750, 2.767) -- (8.800, 2.754) -- (8.850, 2.737) -- (8.900, 2.718) -- (8.950, 2.696) -- (9.000, 2.672) -- (9.050, 2.645) -- (9.100, 2.615) -- (9.150, 2.584) -- (9.200, 2.550) -- (9.250, 2.514) -- (9.300, 2.476) -- (9.350, 2.436) -- (9.400, 2.394) -- (9.450, 2.350) -- (9.500, 2.305) -- (9.550, 2.257) -- (9.600, 2.208) -- (9.650, 2.157) -- (9.700, 2.104) -- (9.750, 2.050) -- (9.800, 1.993) -- (9.850, 1.935) -- (9.900, 1.875) -- (9.950, 1.814) -- (10.000, 1.750) -- (10.050, 1.685) -- (10.100, 1.618) -- (10.150, 1.549) -- (10.200, 1.477) -- (10.250, 1.404) -- (10.300, 1.329) -- (10.350, 1.251) -- (10.400, 1.172) -- (10.450, 1.090) -- (10.500, 1.005) -- (10.550, 0.918) -- (10.600, 0.828) -- (10.650, 0.736) -- (10.700, 0.640) -- (10.750, 0.542) -- (10.800, 0.440) -- (10.850, 0.336) -- (10.900, 0.227) -- (10.950, 0.116) -- (11.000, 0.000);
		\fill[purple] (0.000, 0.000) -- (0.050, 0.202) -- (0.100, 0.376) -- (0.150, 0.525) -- (0.200, 0.651) -- (0.250, 0.754) -- (0.300, 0.838) -- (0.350, 0.903) -- (0.400, 0.951) -- (0.450, 0.983) -- (0.500, 1.002) -- (0.550, 1.008) -- (0.600, 1.003) -- (0.650, 0.988) -- (0.700, 0.965) -- (0.750, 0.934) -- (0.800, 0.897) -- (0.850, 0.855) -- (0.900, 0.809) -- (0.950, 0.760) -- (1.000, 0.710) -- (1.050, 0.658) -- (1.100, 0.606) -- (1.150, 0.554) -- (1.200, 0.504) -- (1.250, 0.456) -- (1.300, 0.411) -- (1.350, 0.369) -- (1.400, 0.331) -- (1.450, 0.297) -- (1.500, 0.269) -- (1.550, 0.245) -- (1.600, 0.227) -- (1.650, 0.215) -- (1.700, 0.210) -- (1.750, 0.210) -- (1.800, 0.217) -- (1.850, 0.231) -- (1.900, 0.251) -- (1.950, 0.278) -- (2.000, 0.312) -- (2.050, 0.352) -- (2.100, 0.398) -- (2.150, 0.450) -- (2.200, 0.508) -- (2.250, 0.572) -- (2.300, 0.641) -- (2.350, 0.714) -- (2.400, 0.792) -- (2.450, 0.873) -- (2.500, 0.958) -- (2.550, 1.045) -- (2.600, 1.133) -- (2.650, 1.223) -- (2.700, 1.314) -- (2.750, 1.403) -- (2.800, 1.492) -- (2.850, 1.578) -- (2.900, 1.661) -- (2.950, 1.739) -- (3.000, 1.812) -- (3.050, 1.878) -- (3.100, 1.937) -- (3.150, 1.987) -- (3.200, 2.026) -- (3.250, 2.054) -- (3.300, 2.068) -- (3.350, 2.068) -- (3.400, 2.052) -- (3.450, 2.019) -- (3.500, 1.967) -- (3.550, 1.893) -- (3.600, 1.797) -- (3.650, 1.677) -- (3.700, 1.531) -- (3.750, 1.357) -- (3.800, 1.154) -- (3.850, 0.918) -- (3.900, 0.648) -- (3.950, 0.343) -- (4.000, 0.000);
		\fill[purple] (7.000, 0.000) -- (7.050, 0.107) -- (7.100, 0.209) -- (7.150, 0.305) -- (7.200, 0.396) -- (7.250, 0.482) -- (7.300, 0.563) -- (7.350, 0.639) -- (7.400, 0.711) -- (7.450, 0.778) -- (7.500, 0.840) -- (7.550, 0.899) -- (7.600, 0.953) -- (7.650, 1.004) -- (7.700, 1.051) -- (7.750, 1.094) -- (7.800, 1.134) -- (7.850, 1.171) -- (7.900, 1.204) -- (7.950, 1.234) -- (8.000, 1.261) -- (8.050, 1.286) -- (8.100, 1.307) -- (8.150, 1.326) -- (8.200, 1.343) -- (8.250, 1.357) -- (8.300, 1.368) -- (8.350, 1.378) -- (8.400, 1.385) -- (8.450, 1.390) -- (8.500, 1.394) -- (8.550, 1.395) -- (8.600, 1.395) -- (8.650, 1.393) -- (8.700, 1.389) -- (8.750, 1.384) -- (8.800, 1.377) -- (8.850, 1.369) -- (8.900, 1.359) -- (8.950, 1.348) -- (9.000, 1.336) -- (9.050, 1.322) -- (9.100, 1.308) -- (9.150, 1.292) -- (9.200, 1.275) -- (9.250, 1.257) -- (9.300, 1.238) -- (9.350, 1.218) -- (9.400, 1.197) -- (9.450, 1.175) -- (9.500, 1.152) -- (9.550, 1.129) -- (9.600, 1.104) -- (9.650, 1.078) -- (9.700, 1.052) -- (9.750, 1.025) -- (9.800, 0.997) -- (9.850, 0.968) -- (9.900, 0.938) -- (9.950, 0.907) -- (10.000, 0.875) -- (10.050, 0.842) -- (10.100, 0.809) -- (10.150, 0.774) -- (10.200, 0.739) -- (10.250, 0.702) -- (10.300, 0.664) -- (10.350, 0.626) -- (10.400, 0.586) -- (10.450, 0.545) -- (10.500, 0.502) -- (10.550, 0.459) -- (10.600, 0.414) -- (10.650, 0.368) -- (10.700, 0.320) -- (10.750, 0.271) -- (10.800, 0.220) -- (10.850, 0.168) -- (10.900, 0.114) -- (10.950, 0.058) -- (11.000, 0.000);
	\end{tikzpicture}
	\begin{tikzpicture}[scale=0.85]
		\node at (5.5, 4.5) {\large Interpolation de Wasserstein};
		\node[red] at (0.5, 2.3) {$\mu$};
		\node[blue] at (9.5, 2.7) {$\nu$};
		\fill[red] (0.000, 0.000) -- (0.050, 0.403) -- (0.100, 0.752) -- (0.150, 1.051) -- (0.200, 1.302) -- (0.250, 1.509) -- (0.300, 1.676) -- (0.350, 1.806) -- (0.400, 1.902) -- (0.450, 1.967) -- (0.500, 2.004) -- (0.550, 2.016) -- (0.600, 2.006) -- (0.650, 1.976) -- (0.700, 1.929) -- (0.750, 1.868) -- (0.800, 1.794) -- (0.850, 1.710) -- (0.900, 1.618) -- (0.950, 1.521) -- (1.000, 1.419) -- (1.050, 1.315) -- (1.100, 1.211) -- (1.150, 1.108) -- (1.200, 1.008) -- (1.250, 0.912) -- (1.300, 0.822) -- (1.350, 0.738) -- (1.400, 0.662) -- (1.450, 0.595) -- (1.500, 0.537) -- (1.550, 0.490) -- (1.600, 0.455) -- (1.650, 0.431) -- (1.700, 0.419) -- (1.750, 0.420) -- (1.800, 0.435) -- (1.850, 0.462) -- (1.900, 0.503) -- (1.950, 0.556) -- (2.000, 0.623) -- (2.050, 0.703) -- (2.100, 0.796) -- (2.150, 0.900) -- (2.200, 1.017) -- (2.250, 1.144) -- (2.300, 1.281) -- (2.350, 1.428) -- (2.400, 1.584) -- (2.450, 1.746) -- (2.500, 1.915) -- (2.550, 2.089) -- (2.600, 2.267) -- (2.650, 2.447) -- (2.700, 2.627) -- (2.750, 2.807) -- (2.800, 2.984) -- (2.850, 3.156) -- (2.900, 3.321) -- (2.950, 3.478) -- (3.000, 3.624) -- (3.050, 3.757) -- (3.100, 3.874) -- (3.150, 3.973) -- (3.200, 4.052) -- (3.250, 4.107) -- (3.300, 4.136) -- (3.350, 4.137) -- (3.400, 4.105) -- (3.450, 4.038) -- (3.500, 3.933) -- (3.550, 3.787) -- (3.600, 3.595) -- (3.650, 3.355) -- (3.700, 3.063) -- (3.750, 2.715) -- (3.800, 2.307) -- (3.850, 1.836) -- (3.900, 1.297) -- (3.950, 0.686) -- (4.000, 0.000);
		\fill[blue] (7.000, 0.000) -- (7.050, 0.215) -- (7.100, 0.418) -- (7.150, 0.611) -- (7.200, 0.792) -- (7.250, 0.964) -- (7.300, 1.126) -- (7.350, 1.278) -- (7.400, 1.421) -- (7.450, 1.555) -- (7.500, 1.681) -- (7.550, 1.798) -- (7.600, 1.907) -- (7.650, 2.008) -- (7.700, 2.102) -- (7.750, 2.188) -- (7.800, 2.268) -- (7.850, 2.341) -- (7.900, 2.408) -- (7.950, 2.468) -- (8.000, 2.522) -- (8.050, 2.571) -- (8.100, 2.614) -- (8.150, 2.652) -- (8.200, 2.685) -- (8.250, 2.713) -- (8.300, 2.737) -- (8.350, 2.755) -- (8.400, 2.770) -- (8.450, 2.781) -- (8.500, 2.787) -- (8.550, 2.790) -- (8.600, 2.790) -- (8.650, 2.785) -- (8.700, 2.778) -- (8.750, 2.767) -- (8.800, 2.754) -- (8.850, 2.737) -- (8.900, 2.718) -- (8.950, 2.696) -- (9.000, 2.672) -- (9.050, 2.645) -- (9.100, 2.615) -- (9.150, 2.584) -- (9.200, 2.550) -- (9.250, 2.514) -- (9.300, 2.476) -- (9.350, 2.436) -- (9.400, 2.394) -- (9.450, 2.350) -- (9.500, 2.305) -- (9.550, 2.257) -- (9.600, 2.208) -- (9.650, 2.157) -- (9.700, 2.104) -- (9.750, 2.050) -- (9.800, 1.993) -- (9.850, 1.935) -- (9.900, 1.875) -- (9.950, 1.814) -- (10.000, 1.750) -- (10.050, 1.685) -- (10.100, 1.618) -- (10.150, 1.549) -- (10.200, 1.477) -- (10.250, 1.404) -- (10.300, 1.329) -- (10.350, 1.251) -- (10.400, 1.172) -- (10.450, 1.090) -- (10.500, 1.005) -- (10.550, 0.918) -- (10.600, 0.828) -- (10.650, 0.736) -- (10.700, 0.640) -- (10.750, 0.542) -- (10.800, 0.440) -- (10.850, 0.336) -- (10.900, 0.227) -- (10.950, 0.116) -- (11.000, 0.000);
		\fill[purple] (3.500, 0.000) -- (3.550, 0.279) -- (3.600, 0.526) -- (3.650, 0.758) -- (3.700, 0.977) -- (3.750, 1.171) -- (3.800, 1.343) -- (3.850, 1.500) -- (3.900, 1.643) -- (3.950, 1.768) -- (4.000, 1.874) -- (4.050, 1.965) -- (4.100, 2.041) -- (4.150, 2.103) -- (4.200, 2.151) -- (4.250, 2.184) -- (4.300, 2.202) -- (4.350, 2.206) -- (4.400, 2.196) -- (4.450, 2.172) -- (4.500, 2.133) -- (4.550, 2.078) -- (4.600, 2.009) -- (4.650, 1.923) -- (4.700, 1.819) -- (4.750, 1.696) -- (4.800, 1.582) -- (4.850, 1.413) -- (4.900, 1.279) -- (4.950, 1.102) -- (5.000, 1.030) -- (5.050, 0.804) -- (5.100, 0.828) -- (5.150, 0.838) -- (5.200, 0.760) -- (5.250, 0.925) -- (5.300, 1.088) -- (5.350, 1.157) -- (5.400, 1.422) -- (5.450, 1.538) -- (5.500, 1.746) -- (5.550, 1.912) -- (5.600, 2.060) -- (5.650, 2.209) -- (5.700, 2.342) -- (5.750, 2.459) -- (5.800, 2.564) -- (5.850, 2.656) -- (5.900, 2.736) -- (5.950, 2.805) -- (6.000, 2.862) -- (6.050, 2.908) -- (6.100, 2.944) -- (6.150, 2.969) -- (6.200, 2.985) -- (6.250, 2.989) -- (6.300, 2.984) -- (6.350, 2.970) -- (6.400, 2.948) -- (6.450, 2.919) -- (6.500, 2.874) -- (6.550, 2.821) -- (6.600, 2.761) -- (6.650, 2.700) -- (6.700, 2.620) -- (6.750, 2.525) -- (6.800, 2.430) -- (6.850, 2.336) -- (6.900, 2.211) -- (6.950, 2.080) -- (7.000, 1.955) -- (7.050, 1.814) -- (7.100, 1.644) -- (7.150, 1.485) -- (7.200, 1.317) -- (7.250, 1.113) -- (7.300, 0.921) -- (7.350, 0.715) -- (7.400, 0.477) -- (7.450, 0.255) -- (7.500, 0.000);
	\end{tikzpicture}
\end{center}

We recall that the computation of the barycenter $\mu$ of the $N$ distributions
$\nu_i$ associated to the weights $\lambda_i$ boils down to the following
minimization problem.
$$ \min_{\mu \in \mathcal{P}(\Omega)} \sum_{i=1}^N \lambda_i W_p^p(\mu, \nu_i) $$
Now observe what the barycenter looks like with $4$ distributions:
\begin{center}
	\includegraphics[scale=0.4]{wasserstein_bar.png}
\end{center}

\sect{How to compute Optimal Transport?}

\subs{Duality}

To do the computations, we embrace the setting of discrete spaces. Our
probability mesures are sum of diracs impulses:

$$ \mu = \sum_{i=1}^n a_i \delta_{x_i} \qquad \nu = \sum_{j =1}^m b_j \delta_{y_j} $$
Our cost function is a matrix $D \in \R_+^{n \times m}$ such that $D_{ij} = d(x_i, y_j)^p$. Our joint distribution $P$ is also a matrix of $\R_+^{n \times m}$ that has to satisfy several constraints that can be represented matricially. The set which $P$ belongs to is:

$$ U(a, b) = \left\{ P \in \R_+^{n \times m}~|~P \mathbbm{1}_m = a, \, P^\trans \mathbbm{1}_n = b \right\} $$

\DEF{
    The Wasserstein distance in a discrete space is defined as:
	$$ W_p^p(\mu, \nu) = \min_{P \in U(a, b)} \langle P, D \rangle $$
	\vspace{-6mm}
}

\paragraph{Dual} We now focus on the dual problem. For this we introduce the lagrangian.

$$ L(P, \alpha, \beta) = \langle P, D \rangle + \alpha^\trans (a - P \mathbbm{1}_m) + \beta^\trans (b - P^\trans \mathbbm{1}_n) $$
The objective function of the dual is:
$$ g(\alpha, \beta) = \min_{P \geqslant 0} L(P, \alpha, \beta) = \min_{P \geqslant 0} \alpha^\trans a + \beta^\trans b + \langle P, D - \alpha \mathbbm{1}_m^\trans - \mathbbm{1}_n \beta^\trans \rangle $$
If the matrix $D - \alpha \mathbbm{1}_m^\trans - \mathbbm{1}_n \beta^\trans$ has a negative coefficient, then by having this coefficient grow toward $+\infty$ we obtain a minimum of $-\infty$. But if all the coefficients are non negative, then  $\langle P, D - \alpha
\mathbbm{1}_m^\trans - \mathbbm{1}_n \beta^\trans \rangle$ is non negative and we just have to take $P=0$ to obtain a non zero value. We therefore obtain:
$$ g(\alpha, \beta) = \left\{ \begin{array}{ll}
\alpha^\trans a + \beta^\trans b & \text{si } D - \alpha \mathbbm{1}_m^\trans - \mathbbm{1}_n \beta^\trans \geqslant 0 \\
-\infty & \text{sinon}
\end{array}\right. $$

Finally, the dual problem is the maximization of this function $g$.

\DEF{
	The \textbf{dual} of the optimal transport problem is:
	$$ W_p^p(\mu, \nu) = \max_{\alpha \in \R^n, \, \beta \in \R^m \atop \alpha_i + \beta_j \leqslant D(x_i, y_j)^p} \alpha^\trans a + \beta^\trans b $$
	\vspace{-5mm}
}

Using a minimum flow solver, we can solve this linear optimization problem in
$\mathcal{O}(n^3 \log (n))$. But the solution is not stable, as illustrated
on the following figure. If we slightly move our distance $D$ or if we change
the source and destination mesures, the value of $P$ can drastically change,
since $P$ is a vertex of our simplex. One solution is to regularize the problem to fall back on a set $U_\alpha(a, b)$
strictly convex. This set is strictly contained in $U(a, b)$ and the optimal solution of this new problem won't be optimal for the initial problem, but will remain valid.
Finally, the regularization that we will describe allows to obtain an
interative algorithm of small complexity.
\begin{center}
	\begin{tikzpicture}[scale=20, >={latex}]
		\clip (0.2, -0.01) rectangle (0.51, 0.2);
		\draw[blue, thick] (0.1420, 0.1407) -- (0.5970, -0.1407);
		\draw[black, ->, thick] (0.2038, 0.1216) -- (0.2507, 0.1975);
		\fill[brown!40] (0.4503, 0.1444) -- (0.2251, 0.1444) -- (0.3695, 0.0000) -- (0.4503, -0.0000);
		\fill[greenTikz!60] (0.3827, 0.0946) -- (0.3827, 0.1104) -- (0.3804, 0.1148) -- (0.3759, 0.1206) -- (0.3692, 0.1263) -- (0.3602, 0.1307) -- (0.3556, 0.1321) -- (0.3489, 0.1336) -- (0.3264, 0.1336) -- (0.3219, 0.1321) -- (0.3151, 0.1278) -- (0.3129, 0.1234) -- (0.3129, 0.1191) -- (0.3151, 0.1133) -- (0.3241, 0.1018) -- (0.3444, 0.0888) -- (0.3534, 0.0859) -- (0.3714, 0.0859) -- (0.3804, 0.0917);
		\node[blue] at (0.3444, 0.0888) {$\bullet$};
		\node[blue, below=2] at (0.3444, 0.0888) {$P_\alpha$};
		\node[greenTikz] at (0.3489, 0.1119) {$\bullet$};
		\node[red] at (0.3695, 0.0000) {$\bullet$};
		\node[red, above=2] at (0.3695, 0) {$P$};
		\node at (0.482, 0.08) {$U(a, b)$};
		\node at (0.413, 0.12) {$U_\alpha(a, b)$};
		\node at (0.213, 0.165) {$D$};
	\end{tikzpicture}
\end{center}

\subs{Regularization}

The regularization that we choose consist in maximizing the entropy of the
joint-probability $P$. We exhibit a new parameter $\gamma$ that controls
the regularization and we obtain the following distance:

\DEF{
	The \textbf{regularized Wasserstein distance} is defined as follows:
	$$ W_\gamma(\mu, \nu) = \min_{P \in U(a, b)} \langle P, D \rangle - \gamma H(P) $$
	\vspace{-5mm}
}

\paragraph{Remark}
Recall that we can consider two random variables $X$ and $Y$ such that $X \sim \mu$, $Y \sim \nu$ and $X,Y \sim P$. The entropy of $P$ is maximum when $X$ and $Y$ are independant, \emph{ie.} when $P = ab^\trans$ and this maximum entropy is thus $H(\mu) + H(\nu)$. 
The regularization therefore gives a matrix whose positive values are more
distributed.

\begin{center}
	\includegraphics[scale=0.25]{reg_wass.png}
\end{center}

By strict convexity, there exist a unique matrix $P_\gamma$ that minimizes the distance:
$$ P_\gamma = \argmin_{P \in U(a, b)} \langle P, D \rangle - \gamma H(P) $$

\PROP{
    There exist a unique couple of vectors $u$ and $v$ belonging to
    $\R_+^n$ and $\R_+^m$ such that :
	$$ P_\gamma = \text{diag}(u) K \text{diag}(v), \qquad K = e^{-D / \gamma} $$
	\vspace{-6mm}
}

\dem

We write the new laplacian:

$$ L(P, \alpha, \beta) = \sum_{ij} \left( P_{ij} D_{ij} + \gamma P_{ij} \left( \log_2 P_{ij} - 1 \right) \right) + \alpha^\trans (a - P \mathbbm{1}_m) + \beta^\trans (b - P^\trans \mathbbm{1}_n) $$
We compute the partial derivative with respect to $P_{ij}$:
$$ \dfrac{\partial L}{\partial P_{ij}} = M_{ij} + \gamma \log_2 P_{ij} - \alpha_i - \beta_j $$
The dual is the maximization of the minimum of the laplacian with respect
to $P$. We must look for $P$ such that the partial derivative is equal to zero:
$$ \dfrac{\partial L}{\partial P_{ij}} = 0 \Rightarrow P_{ij} = e^{\frac{\alpha_i}{\gamma}} e^{-\frac{D_{ij}}{\gamma}}  e^{\frac{\beta_j}{\gamma}} = u_i K_{ij} v_j $$
\findem

\paragraph{Sinkhorn}
We have at our disposition an interative method to find this matrix $P_\gamma$. We use the condition $P_\gamma \in U(a,b)$:
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
\text{diag}(u) K \text{diag}(v) \mathbbm{1}_m & = & a \\
\text{diag}(v) K^\trans \text{diag}(u) \mathbbm{1}_n & = & b
\end{array} \right. $$
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
\text{diag}(u) K v & = & a \\
\text{diag}(v) K^\trans u & = & b
\end{array} \right. $$
Then, we introduce $\odot$ the coefficient-wise product:
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
u \odot K v & = & a \\
v \odot K^\trans u & = & b
\end{array} \right. $$
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrc}
u & = & a \, / \, K v \\
v & = & b \, / \, K^\trans u
\end{array} \right. $$
Sinkhorn's algorithm is thus the following: \\
\begin{algorithm}[H]
	\caption{Sinkhorn}
	\Repeat{convergence of $u, v$}{
		$u \gets a \, / \, K v$ \;
		$v \gets b \, / \, K^\trans u$ \;
	}
\end{algorithm}

\paragraph{Complexity}
It has been proved that the algorithm converges in linear time. Moreover,
an iteration is done in $\mathcal{O}(mn)$ but the computation can be
parallelized. It is also possible to use grid convolution to obtain a
complexity of $\mathcal{O}(n \log n)$.

\paragraph{Dual}
On peut aussi relever la formulation du dual avec régularisation :
$$ W_\gamma(\mu, \nu) = \max_{\alpha, \beta} \alpha^\trans a + \beta^\trans b - \gamma \left( e^{\alpha / \gamma} \right)^\trans K \left( e^{\beta / \gamma} \right) $$

\paragraph{Lien avec la divergence KL}
We write the kernel:
$$ K_\gamma(i, j) = \exp \left( - \dfrac{d_{i, j}}{\gamma} \right) $$
We obtain a new formulation of the entropic regularization with the KL-divergence:
$$ W_\gamma(\mu, \nu) = \min_{P \in U(a, b)} \gamma KL(P | K_\gamma) $$

\sect{Optimal Transport and Machine Learning}

\subs{Compute the optimal transport of what?}

Since Sinkhorn can be parallelized, one can compute a large amount of Optimal
Transport distances: for instance look into a database for similarity based
retrieval, blablabla

\paragraph{Color histograms}

For image retrieval blablabla...

\paragraph{Cloud of word}

For similar documents retrieval blabla

\subs{Wassersteinization}

Introduce Wasserstein distance into optimizations problem.

We show that we can computationaly differentiate, and therefore we can deal
with it in a variety of optimization problem, including machine learning
related opt. problems.

Somehow revisiting machine learning problems.

\paragraph{Averaging data}

histograms, distribution, images, experiments (repetition of same expriment,
MEG mesures, average over the experiments might be relevant)

\paragraph{Aggregating distributions}

From a very big dataset, split it in $J$ parts, learn the distribution (eg. w/
Bayesaian Learning) from each of the part in parralel (on $J$ machines):
the $J$ distributions should not be very different, but still the variance in
the data will cause some variation: the idea is to use Wasserstein averaging
to aggregate the distributions, and get a more meaningful one. Theoretic
guaranties! :D

Wasserstein posterior

%\subs{Semi-suppervised learning setting}

\paragraph{Semi-supervised learning setting}

Wasserstein propagation: smooth a graph $(V,E)$ where an histogram is associated
$\mu_v$ to each vertex $v$, with a susbset $S$ of the vertices that have a
fixed histogram (known data).

Solve the following optimization problem.

$$
    \underset{\substack{\mu_i \in \mathcal{P}\left(\Omega\right) \\ \text{for }i \in V\setminus S}}\min \sum_{(e_1, e_2) \in E} W_2^2\left(\mu_{e_1}, \mu_{e_2}\right)
$$

blabla

\paragraph{omg dictionnary learning what is this????????? :oooooo}

\subs{Wasserstein PCA, Geodesics and Inverse problem}

Since the space of distributions with the OT geometry has negative curvature,
we generalize the PCA (ie. projecting on one dimension spaces) into Generalized
Principal Geodesics. We then want to explain stuff out of what we found;
eg. affaire Bonneel vs nounours