### 2019-10-27 report/implementation.tex
Rework some sentences and notations

\chapter{Implementation}

We decided to implement the last paper to compute Wasserstein distances
on images in an efficient way. We then use convolutional Sinkhorn in an SVM
classifier as Marco Cuturi used Sinkhorn in his paper to classify images of
the MNIST digit database \cite{Cut}. The implementation of the convolutional
Wassertein distance has been written in C++. We then write the Gramm matrix
of a subset of the MNIST database in a file and use it in a Python script
to use SVM from Sklearn library.

\paragraph{How to use our code}
OpenMP has been used to parallelize the code. If you don't have OpenMP already
installed, you can install it with \verb|sudo apt install libomp-dev|.
To compile the C++ code you just have to type \verb|make|. Then you need
to obtain the train image datatset and the train label dataset of the MNIST
database \url{http://yann.lecun.com/exdb/mnist/}. To do so you have to run
the script \verb|get_mnist.sh| which will download and rename the files
correctly. The executable \verb|main| created by \verb|make| command can be
used in four ways.

\begin{itemize}
	\item  \verb|./main bar| to reproduce an image similar to the
	\figurename~\ref{barp}; An image containing different barycenters
	of four shapes.

	\item \verb|./main card| to reproduce images similar to the
	\figurename~\ref{cardp}; An interpolation of two images of cards.

	\item \verb|./main wass > kernel.txt| to compute a kernel with
	the convolutional Wasserstein distance of \verb|N_TRAIN| images of
	the MNIST dataset. We project \verb|N_TEST| other digit images of
	the MNIST dataset on the kernel space created. You can change the
	parameter \verb|N_TRAIN| and \verb|N_TEST| at the beginning of the
	file \verb|src/main.cpp|. The kernel will then be written if the
	file \verb|kernel.txt|.

	\item \verb|./main eucl > kernel.txt| to do the same as the previous
	command but with Euclidean distance instead of Wasserstein distance.
\end{itemize}
Finally when a kernel file is computed you can test SVM on it
thanks to the python script \verb|learning.py|. Because computing
Wasserstein kernel may take a while to compute we provide a pre-computed
Wasserstein kernel \verb|kernels/wass.txt| as well as an Euclidean kernel
\verb|kernels/eucl.txt|. You can then run:
\begin{center}
	\verb=python3 learning.py < [kernels/wass.txt | kernels/eucl.txt]=
\end{center}
It will print the performances of these two kernels, show the projection of
the data on the two principle components and finally show the kernel matrix.

\paragraph{Barycenters}
We first computed some barycenters to be sure that everything works. The
code for computing Wasserstein distances and barycenters is in the file
\verb|src/wasserstein.cpp|. The code that produce the barycenters that we
will present is in the file \verb|examples.cpp|. The first example is the
reproduction of \figurename~\ref{barp}. Here is what we obtained:

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{bars_own.png}
	\caption{Reproduction of \figurename~\ref{barp}}
\end{figure}

The result seems to be correct. We then try on larger images, the card images that we can find in the original paper presenting convolutional Wassertein \cite{Goe++}. Here is what we obtained:

\begin{figure}[h]
	\centering
	\begin{tikzpicture}
		\node[] at (0, 0) {\includegraphics[scale=0.3]{bar0.png}};
		\node[] at (2.8, 0) {\includegraphics[scale=0.3]{bar3.png}};
		\node[] at (5.6, 0) {\includegraphics[scale=0.3]{bar6.png}};
		\node[] at (8.4, 0) {\includegraphics[scale=0.3]{bar8.png}};
		\node[] at (11.2, 0) {\includegraphics[scale=0.3]{bar11.png}};
		\node[] at (14, 0) {\includegraphics[scale=0.3]{bar14.png}};
	\end{tikzpicture}
	\caption{Reproduction of \figurename~\ref{cardp}}
\end{figure}

\paragraph{SVM usage}
As proposed in the paper of Cuturi \cite{Cut}, we used the Wasserstein distance to compute the Gram matrix for 500 image and we used it as a kernel in a SVM. On a test set of 1500 images we obtained 92.4\% of good answers. Whereas with a kernel using Euclidean distance we obtain a good answer on only 88.4\%. To check these results you can run the python script \verb|learning.py| on the two kernels that are in the \verb|kernels| folder.