### 2019-10-27 report/papers.tex

\chapter{Review of papers}

\myminitoc

\sect{Sinkhorn Distances: Lightspeed Computation of Optimal Transport \cite{Cut}}

In this paper published in 2013, Marco Cuturi proposed one of the currently
most used algorithm to compute Wasserstein distances. He smoothed the classic
optimal transport problem thanks to an entropic regularization and it leads
to a new notion of distance which has a faster solver. This follows the
regularization idea the we introduced in the first chapter. He also shows
in this paper that using such a regularization gives better results on
MNIST classification problems than using the standard Wasserstein distance
(or obviously using the Euclidean distance). This paper is part of the search
for adequate and tractable distances in machine learning. Wasserstein distance
is more suited than Euclidean distance to many problems in machine learning,
but because the complexity to compute it is $\mathcal{O}(n^3 \log n)$ when
$n$ is the size of the data, Wasserstein distance is rarely used. That is
why such techniques to compute the Wasserstein distance had to be developped.

\paragraph{Idea}
The main ideas of this paper are the use of an entropic-regularization to
simplify computation when using Sinkhorn iterative algorithm and the idea
of computing the distances between one point and a family of other points in
the same time, to use matrix-matrix multiplications instead of matrix-vector
multiplications while computing distances between pair of points. Thus the
algorithm can be implemented on GPU architectures.

\paragraph{Regularization}
We recall that the goal of the optimal transport problem in the Kantorovich
formulation is to find a joint distribution $P \in \R^{n \times n}$ where
the marginals equal the two distributions for which we want to compute the
distance. That is to say $P \mathbf{1}_n = a$ and $P^\trans \mathbf{1}_n = b$
when we compute the distance between $a$ and $b$. Following the notations
introduced in the first chapter, we call $U(a, b)$ the set of all possible
values for $P$. By denoting the entropy by $H$ we can show that:

$$ H(P) \leqslant H(a) + H(b) $$

And $H(P) = H(a) + H(b)$ when $P = a b^\trans$ (that is, when $a$ and $b$ are independant).
We then restrict $P$ to the new set:

$$ U_\alpha(a, b) = \{ P \in U(a, b)
\, | \, \text{KL}(P || a b^\trans) \leqslant \alpha \} = \{ P \in U(a, b)
\, | \, H(P) \geqslant H(a) + H(b) - \alpha \} $$

Two reasons are given to this restriction. The first one is that it will be
easier to compute the distance. The second is that without regularization
$P$ has almost $2n-1$ non-zero coefficients. Thus the transport is almost
deterministic and it is not natural. That is why smoothing the transport
plan with entropic regularization is presented as a good idea. In the paper,
it is shown that this restriction induce a new distance:

$$ d_{M, \alpha} = \min_{P \in U_\alpha(a,b)} \langle P, M \rangle $$

Where $M$ is a distance matrix.

\paragraph{Computation}
By duality theory it is obtained that to $\alpha$, corresponds a value
$\lambda \geqslant 0$ such that the distance $d_{M, \alpha}$ is equal to
the distance $d_M^\lambda$ defined by:
$$ d_M^\lambda(a, b) = \langle P^\lambda, M \rangle, \quad \text{where } P^\lambda = \argmin_{P \in U(a, b)} \langle P, M \rangle - \dfrac{1}{\lambda} H(P) $$

Then Cuturi shows that this new distance can be computed with a much cheaper
cost than the original Wassertein distance $d_M$. In fact the iterative
Sinkhorn algorithm presented below converge in few steps to the transport
plan $P^\lambda$. The \figurename~\ref{recap1} summarize the relationships
between all these distances.

\vspace{3mm}

\begin{algorithm}[H]
	\caption{\textsc{Sinkhorn}$\left( a, \{ b_i \}_{i=1}^N, M, \lambda \right)$}
	$B \gets \left[ b_1, \dots, b_N \right] \in \R^{n \times N}$ \;
	$K \gets \exp \left( \lambda M \right)$ \;
	$u = \left[ u_1, \dots, u_N \right] \gets \mathbbm{1}_{n \times N} \, / \, n$ \;
	\Repeat{convergence of $u$}{
		$u \gets a \, / \, \left( K \left( B \, / \, \left( K^\trans u \right) \right) \right)$ \;
	}
	$v = \left[ v_1, \dots, v_N \right] \gets B \, / \, \left( K^\trans u \right)$ \;
	$P^\lambda_i \gets \text{diag}(u_i) K \text{diag}(v_i)$ \;
	\Return $d = \left[ d_M^\lambda(a, b_1), \dots, d_M^\lambda(a, b_N) \right] = \left[ \left\langle M, P^\lambda_1 \right\rangle, \dots, \left\langle M, P^\lambda_N \right\rangle \right]$
\end{algorithm}
\vspace{3mm}

\begin{figure}
	\centering
	\includegraphics[scale=0.5]{recap1.png}
	\captionsetup{justification=centering}
	\caption{Graphical representation of different distances and their relationships.}
	\label{recap1}
\end{figure}

\paragraph{Results}
The regularized Wasserstein distance is then used in a SVM to classifiy
images of the MNIST database. The \figurename~\ref{comp} compare the error
on a test set for different distances used in the SVM. The distance EMD
is the classical optimal transport. As we can see the regularized version
beats all other distances. Furthermore he evaluated the execution time by
stopping the iterations when $\| d_{i+1} / d_i - 1 \| < 10^{-4}$ where $d_i$
is the distance obtained at the end of the iteration $i$. He gave a comparison
between EMD and regularized Wasserstein for different values of $\lambda$
in the \figurename~\ref{comp2}, highlighting the efficiency of computing
the regularized optimal transport.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.55]{comp.png}
	\caption{Test error for different distances}
	\label{comp}
\end{figure}
\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{comp2.png}
\caption{Execution time for different distances}
\label{comp2}
\end{figure}

\newpage

\sect{Wasserstein GAN \cite{Arj++}}

This paper proposes a new GAN training algorithm based on Wasserstein
distance. When we learn a probability distribution we need to have a notion of
divergence between probability distributions. In a first time they show the
advantages of the Wasserstein distance over all well-known other divergences
on probabilities. Thanks to their WGAN (for Wasserstein GAN), they cure the
problem of the need of maintaining a balance between training discriminator
and training generator. They also observed that the mode dropping phenomenon
is reduced. Furthermore their GAN allows to plot the Wasserstein distance
continuously which is very usefull for debugging and having an estimation
of the quality.

\paragraph{Advantage of Wasserstein distance}
They illustrate the interest of Wasserstein distance with a toy problem,
that is the convergence of the probability $\Pp_\theta = g_\theta(Z) =
(\theta, Z)$ toward $\Pp_0 = (0, Z)$ for $Z \sim U([0, 1])$ when $\theta$
tends to 0. We have:
\begin{itemize}
	\item $W(\Pp_0, \Pp_\theta) = | \theta |$ where $W$ is the Wasserstein distance.
	\item $JS(\Pp_0, \Pp_\theta) = \left\{ \begin{array}{ll}
			\log 2 & \text{if } \theta \neq 0 \\
			0 & \text{if } \theta = 0
		\end{array} \right.$ where $JS$ is the Jensen-Shannon divergence.
	\item $KL(\Pp_0, \Pp_\theta) = JS(\Pp_\theta, \Pp_0) = \left\{ \begin{array}{ll}
			+ \infty & \text{if } \theta \neq 0 \\
			0 & \text{if } \theta = 0
		\end{array} \right.$ where $KL$ is the Kullback-Leibler divergence.
	\item $\delta(\Pp_0, \Pp_\theta) = \left\{ \begin{array}{ll}
			1 & \text{if } \theta \neq 0 \\
			0 & \text{if } \theta = 0
		\end{array} \right.$ where $\delta$ is the total variation distance defined as \\ $\delta(\Pp_0, \Pp_1) = \sup_{A \in \Sigma} | \Pp_0(A) - \Pp_1(A) |$ with $\Sigma$ the set of all Borel subsets.
\end{itemize}

Furthermore the Wasserstein distance is the only one for which there is
convergence. Then a gradient descent can be applied using the Wasserstein
distance while it cannot be applied with other divergences. This is a just an
example but they also proved the following theorem which shows the continuity
under some assumptions.

\PROP{
	Let $\Pp_r$ be a distribution over $\mathcal{X}$ and let $Z$ be a distribution over another space $\mathcal{Z}$ (e.g Gaussian). Let $g : \R^d \times \mathcal{Z} \rightarrow \mathcal{X}$ be a function that will be denoted $g_\theta(z)$ for $g(\theta, z)$. We denote by $\Pp_\theta$ the distribution of $g_\theta(Z)$. Then:
	\begin{itemize}
		\item If $g$ is continuous in $\theta$, so is $W(\Pp_r, \Pp_\theta)$.
		\item If $g$ is locally Lipschitz and satisfies the following regularity assumption: By denoting the local Lipschitz constant $L(\theta, z)$ we have $\E_{z \sim Z} \left[ L(\theta, z) \right] < + \infty$. Then $W(\Pp_r, \Pp_\theta)$ is continuous everywhere and almost differentiable everywhere.
		\item The two previous statements are false for Jensen-Shannon and KL divergences.
	\end{itemize}
}

They also give another theorem that well show that the Wasserstein distance
is a very natural divergence for probability distributions.

\PROP{
	Consider a sequence of distributions $(\Pp_n)$ on a compact space
	$\mathcal{X}$. If $W(\Pp_n, \Pp) \rightarrow 0$ for a certain
	distribution $\Pp$, then $\Pp_n \xrightarrow{\mathcal{D}} \Pp$
	where the limit is in the sense of the convergence in distribution.
}

\paragraph{WGAN}
To construct their GAN they need to compute the gradient of $W(\Pp_r,
\Pp_\theta)$ with respect to $\theta$. For that they consider the following
dual formulation of the Wasserstein distance:

$$ W(\Pp_r ,\Pp_\theta) = \max_{\| f \|_L \leqslant 1} \E_{x \sim \Pp_r} \left[ f(x) \right] - \E_{x \sim \Pp_\theta} \left[ f(x) \right] $$

Where $\| f \|_L$ denote the Lipshitz coefficient of $f$. If $f$ is the
solution to the maximization problem then the gradient is:

$$ \nabla_\theta W(\Pp_r, \Pp_\theta) = - \E_{z \sim Z} \left[ \nabla_\theta f(g_\theta(z)) \right] $$

To find $f$ they train a neural network parameterized with weights $w$
in a compact space $\mathcal{W}$ that lead to a function $f_w$. Then they
can backpropagate $\theta$ using $\E_{z \sim Z} \left[ \nabla_\theta f_w
(g_\theta(z)) \right]$.  To have $w$ in a compact space, they use weight
clamping. Their algorithm is in the \figurename~\ref{algo}.

\begin{figure}[t]
	\centering
	\includegraphics[scale=0.6]{algorithm.png}
	\caption{Algorithm of the WGAN}
	\label{algo}
\end{figure}

\paragraph{Results}
First, the authors set up a small experiment to showcase the difference
between GAN and WGAN (\figurename~\ref{compa}). There are two 1D Gaussian
distributions, blue for real and green for fake. Train a GAN discriminator and
WGAN critic to optimality, then plot their values over the space. The red curve
is the GAN discriminator output, and the cyan curve is the WGAN critic output.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.36]{wgan_res.png}
	\caption{Comparison between a GAN discriminator and the WGAN Critic}
	\label{compa}
\end{figure}

Both identify which distribution is real and which is fake, but the GAN
discriminator does so in a way that makes gradients vanish over most of
the space. In contrast, the weight clamping in WGAN gives a reasonably nice
gradient over the whole space.

Another important point is that the quality of the generation is correlated
to Wasserstein estimation as we can see in the \figurename~\ref{meaning}. We
can also see that the training of the critic function $f$ works well as the
Wasserstein estimator strictly decrease. The gradient of the loss function is
well computed. But this is not a method to compare the quality of different
GANs as the estimator is defined with a constant scaling factor depending
on the critic's architecture. They finally observed an improvement of the
stability.

\begin{figure}
	\centering
	\includegraphics[scale=0.45]{meaning.png}
	\caption{Sample images at different iterations of the training}
	\label{meaning}
\end{figure}

\sect{Convolutional Wasserstein Distances: Efficient Optimal Transportation on Geometric Domains \cite{Goe++}}

Because a single step of Sinkhorn algorithm has a complexity of
$\mathcal{O}(n^2)$, computing the Wassertein distance may take a while in
many cases, even though the number of iterations needed to converge toward
a good minimizer is small. That is why the researchers who wrote this paper
thought about a more effective solution to perform a step of the
algorithm. They found a way to use Gaussian convolution with a $\mathcal{O}(n
\log n)$ complexity instead of a matrix-vector product. More generally they
use a heat kernel. So instead of the convolution it can also be a sparse
pre-factored linear system which can be computed in a complexity smaller
than $\mathcal{O}(n^2)$. Furthermore this approach does not affect the overall
convergence speed of the scheme, which remains linear. Of course this can
not be applied to all the cases of optimal transport, but it can be used
on very common geometric domains, like images or meshes. In addition of the
computation of Wasserstein distance, the authors also propose a way to use
convolution in the computation of Wasserstein barycenters and Wasserstein
propagation.

\paragraph{Using a heat kernel in the entropy-regularization}
We recall what is the entropy-regularized Wasserstein distance between $\mu_0$
and $\mu_1$ on the domain $M$:

$$
    W_{2, \gamma}^2(\mu, \nu)
        = \inf_{\pi \in \Pi(\mu, \nu)} \left[ \int \int_{M\times M} d(x, y)^2 \pi(x, y) dx dy \, - \, \gamma H(\pi) \right]
        = \gamma \left[ 1 + \min_{\pi \in \Pi(\mu,\nu)} \text{KL}(\pi | \mathcal{K}_\gamma) \right]
$$

Where have kept the same notations as the ones used in the first chapter,
that is

$$
    \Pi = \left\{ \pi \in \mathcal{P}(M \times M) \mid \pi(\cdot, M) = \mu_0, \pi(M, \cdot) = \mu_1 \right\}
$$

$$
    H(\pi) = - \int \int_{M \times M} \pi(x, y) \ln \pi(x, y) dx dy
$$

$$
    \mathcal{K}_\gamma = e^{-d(x, y)^2 / \gamma}
$$
$$
    \text{KL}(\pi | \mathcal{K}) = \int \int_{M \times M} \pi(x, y) \left[ \ln \dfrac{\pi(x, y)}{\mathcal{K}(x, y)} - 1 \right] dx dy
$$

This is a strictly convex problem thanks to the entropy. The idea of the paper
is to use heat kernel because in some cases, the solution of the diffusion
equation can be computed in a effective way. We denote by $\mathcal{H}_t(x,
y)$ the diffusion between $x$ and $y$ after a time $t$ in the heat kernel. Just
remember that a heat kernel is a solution of the heat equation:

$$
    \dfrac{\partial f}{\partial t} = \Delta f
$$

One can show the following that the following approximation holds

$$ \mathcal{K}_\gamma \approx \mathcal{H}_{\gamma / 2}(x, y) $$

We won't store $\mathcal{H}$ because it could have a space complexity in
$\mathcal{O}(n^2)$ and we want a lower complexity. But we generally know
a way to apply $\mathcal{H}$ to a vector. In the case of images we apply
Gaussian convolution with $\sigma^2 = \gamma$. In the case of triangle meshes
we associate a weight to faces proportional to their area. We denote by $a$
the vector of those weights (for images of size $n \times m$ we set $a = 1 /
(nm)$). Then we denote by $L$ the cotangent Laplacian and by $D_a$ the diagonal
matrix whose diagonal is $a$. By discretizing the heat equation, we obtain:

$$
    w = \mathcal{H}_t(v) \Leftrightarrow \left( D_a + t L \right) w = v
$$

The linear system can be solved efficiently by pre-computing a sparse
Cholesky factorization.

\paragraph{Algorithm}
With some computations we arrive at

\vspace{3mm}
\begin{algorithm}[H]
	\caption{\textsc{Convolutional-Sinkhorn}($\mu_0, \mu_1, H_t, a$)}
	$v, w \gets 1$ \;
	\Repeat{convergence of $v, w$}{
		$v \gets \mu_0 \oslash H_t(a \otimes w)$ \;
		$w \gets \mu_1 \oslash H_t(a \otimes v)$ \;
	}
	\Return $2t a^\trans \left[ \left( \mu_0 \otimes \ln v \right) + \left( \mu_1 \otimes \ln w \right) \right]$
\end{algorithm}
\vspace{3mm}

Where $\oslash$ and $\otimes$ denote element-wise operations. As with classical
Sinkhorn algorithm we obtain a simple iterative algorithm where, this time,
each step can be computed in $\mathcal{O}(n \log n)$ in many cases, $n$
being the size of the domain $M$.

\paragraph{Barycenter}
An algorithm for computing barycenters is also provided. We won't enter
into details but we give their algorithm for computing the barycenter of
distributions $\mu_i$ associated with weights $\alpha_i$ for $1 \leqslant
i \leqslant k$:

\vspace{3mm}
\begin{algorithm}[H]
	\caption{\textsc{Convolutional-Barycenter}($\{\mu_i\}, \{ \alpha_i \}, H_t, a$)}
	$v_1, \dots, v_k \gets 1$ \;
	$w_1, \dots, w_k \gets 1$ \;
	\Repeat{convergence of $v_i, w_i$}{
		$\mu \gets 1$ \;
		\For{$i = 1, \dots, k$}{
			$w_i \gets \mu_i \oslash H_t(a \otimes v_i)$ \;
			$d_i \gets v_i \otimes H_t(a \otimes w_i)$ \;
			$\mu \gets \mu \otimes d_i^{\alpha_i}$ \;
		}
		\vspace{3mm}
		// Optional \;
		$\mu \gets $ \textsc{Entropic-Sharpening}$\left( \mu, \max_i H(\mu_i) \right)$ \;
		\vspace{3mm}
		\For{$i = 1, \dots, k$}{
			$v_i \gets v_i \otimes \mu \oslash d_i$ \;
		}
	}
	\Return $2t a^\trans \left[ \left( \mu_0 \otimes \ln v \right) + \left( \mu_1 \otimes \ln w \right) \right]$
\end{algorithm}
\vspace{3mm}
Here entropic sharpening allows us to make the entropy of the result $\mu$
smaller than the maximum entropy of all $\mu_i$. To do so if $H(\mu)$
is greater than $\max_i H(\mu_i)$, then we set $\mu$ to $\mu^\beta$ where
$\beta$ is such that $H(\mu^\beta) = \max_i H(\mu_i)$. We can find such a
$\beta$ using Newton-like methods.

\paragraph{Applications}
This has many applications in geometric domains. Applications that are given are:
\begin{itemize}
	\item Shape interpolation (\figurename~\ref{shape})
		\begin{figure}[h]
			\centering
			\includegraphics[scale=0.22]{shape_int.png}
			\caption{Application to shape interpolation}
			\label{shape}
		\end{figure}
	\item BRDF design
	\item Color histogram manipulation
	\item Skeleton layout
	\item Soft maps
\end{itemize}

Finally \figurename~\ref{barp} and \figurename~\ref{cardp} are two images
of barycenters that they have computed. The first one is an image where
we find several barycenters of four images that are in the corners. The
second is an interpolation between two cards. As we tried to reproduce it
in our implementation part we put them here, for comparison sakes:

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.28]{bars_paper.png}
	\caption{Barycenters of four images}
	\label{barp}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.45]{cards_paper.png}
	\caption{Interpolation of two images of cards for times $t = 0, 0.25, 0.5, 0.75, 1$.}
	\label{cardp}
\end{figure}